### 2018-10-29 00_abstract.tex
Initial ShareLaTeX Import

Rearranging objects on a planar surface arises in a variety of
robotic applications, such as product packaging.  Using two arms can improve
efficiency but introduces new computational challenges. This paper
studies the structure of dual-arm rearrangement for synchronous,
monotone tabletop setups and develops an optimal mixed integer model. It then
describes an efficient and scalable algorithm, which first minimizes
the cost of object transfers and then of moves between objects.
This is motivated by the fact that, asymptotically, object transfers
dominate the cost of solutions. Moreover, a lazy strategy minimizes
the number of motion planning calls and results in significant
speedups.  Theoretical arguments support the benefits of using two
arms and indicate that synchronous execution, in which the two arms perform together either transfers or moves, introduces only a small
overhead.  Experiments support these points and show that the
scalable method can quickly compute solutions close to the optimal for
the considered setup.