### 2018-10-29 09_discussion.tex

The current work demonstrates the underlying structure of synchronized dual-arm rearrangement and proposes an \milp formulation, as well as a scalable algorithm \algo that provides fast, high quality solutions. Existing efficient solvers for reductions of the dual-arm problem made \algo effective. Future work will attempt to explore the $k-$arm case and instances of non-monotone rearrangement. The incorporation of manipulation and regrasp reasoning can extend these methods to more cluttered domains.

% It can be shown however that efficient approximate solvers for more than 2 arms would be hard to design.
% Let us assume a $ \rho $-Approx algorithm, $\mathbb{A}$ exists for the problem of rearrangement with 3 arms. Assume a hypothetical manipulator that incurs a constant transit cost on a bounded plane. The problem of an optimal set of object assignments reduce to 3D-Matching. $ \rho $-Approx $\mathbb{A}$ should return a solution to 3DM. Since 3DM is hard to approximate, $\mathbb{A}$ does not exist.\kiril{The connection with 3DM is not immediate, and I don't think that there's enough information provided here in order to make sense of it. I suggest to omit this description, especially if space is limited.}