### 2020-12-17 001_ntp.tex
Updates from Overleaf

\tase{Monotone tabletop rearrangement challenges arise in a variety of automation scenarios including product sorting or packing.  Performing this task with two robotic manipulators introduces the overhead of coordinating them in the shared workspace, as well as an increase in the size of the underling search space. The objective of this work is to study the feasibility of such dual-arm solutions, providing both theoretical bounds, as well as a fast, and approximate solution. The approach leverages an effective algorithmic decomposition of the problem so as to take advantage of efficient motion planners and mixed-integer linear programming solvers. The proposed solution has been evaluated in settings that include delta-robots as well as \textit{7-dof} manipulators. Interesting extensions of this work correspond to studying the case of additional arms, non-monotone and general manipulation scenarios.}